---
title: "GapMinder Analysis"
author: "brc4ja"
date: "10/14/2019"
output: 
  html_document:
    keep_md: yes 
    fig_caption: yes
    toc: yes
---

# Introduction

Here is my first RMarkdown document. 
In the YAML (the red text denoting title, author, etc) under output, enter between the colon and "html_document". Enter again, tab, and type fig_caption: yes; enter again, and type toc: yes. This will allow a figure and a table of contents in the end document.

# Let's Embed Some R Code

We'll write a R chunk that loads the tidyverse package and then read in the Gapminder dataset from the subdirectory of my project. To insert a R chunk, type ctrl + alt + i. 

```{r}
library(tidyverse)
gm <- read_csv("gapminder.csv")
```


# Investigate the Dataset

Let's take a look at the gm dataset. 

```{r}
head(gm)
```

We have a tibble with country, continent, year, life expectancy, population, and GDP per capita.


# Our First Plot

Showing GDP/capita on X and life expectancy on Y. Specify the figure caption by typing fig.cap="Title" inside the {}. This is referred to as setting the R chunk caption. 

```{r, fig.cap="Life Expectancy vs GDP"}
ggplot(gm, aes(x=gdpPercap, y=lifeExp)) +
  geom_point()
```


# A Better Plot
```{r, fig.cap="Life Expectancy vs GDP, with color"}
ggplot(gm, aes(x=gdpPercap, y=lifeExp)) +
  geom_point(aes(col=continent))+
  scale_x_log10()
```


# Chunk Options

There are many chunk options. Check them out here: https://yihui.name/knitr/options/
Today, we're most interested in the following:
- echo: TRUE by default; will tell chunk whether to include the code in the output. This is great if you don't want to show all the analysis behind your plot.
- results
  - =hide: results will be hidden
  - =hold: will hold results until the end of the R chunk. This is great if you want to have ALL the code, then ALL the outputs as opposed to code/output, code/output, etc
- include: TRUE by default; if FALSE, then the code will not be run. For example, if you have a View() in your RMarkdown, this allows you to hide the View function in your RMarkdown while keeping it in the code so that it can be analyzed by an additional user
- fig.width, fig.height: used to set dimensions in figure output
- cache: FALSE by default; if TRUE, this saves the results of a given piece of code for the next time you knit it. This is helpful if you're making minute changes and want to re-knit the document several times, but the knitting takes a long time. This way, each chunk will be evaluated for changes since the last time the knitting was performed, and if there are no changes, will just use the last run instead of recomputing. 


# Changing Dimensions
```{r, fig.cap="Life Expectancy vs GDP, with color, fig.width=2"}
ggplot(gm, aes(x=gdpPercap, y=lifeExp)) +
  geom_point(aes(col=continent))+
  scale_x_log10()
```
  
  
# Writing an R Markdown Table

See some data using head (which is nicely formatted)

```{r}
head(gm)
```

Now make it into a table.
```{r}
library(knitr)
kable(head(gm))
```

If you like tables, look at **gt** table package. (I made this bold by surrounding the word with two asterisks on each side. You can make something italic by surrounding it with one asterisk on each side.) gt uses dplyr. 


# Changing from html to pdf

In the YAML output, change "html" to "pdf". If this doesn't work, there is an error in LaTeX installation, which we'll come back to later. 

GitHub has difficulty publishing html documents. You need a .md file. Go up to the YAML, and under "html_document:" enter "keep_md: yes"


# Now We Knit

Under the title of this script, click Knit > Knit to html. Alternatively, use control + shift + k.

# Let's Share by Using GitHub

Make a new repository in GitHub, leaving R for now. 